---
title: "CreateFlareData"
output: pdf_document
date: "2023-05-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Write function that takes in the two nc files (one for 1-min flux, the other one for identified flares), and returns the dataset of 1-min peak-fluxes, with an additional column whether a flare was identified.
```{r}
library(ncdf4)
library(dplyr)
library(readr)
library(RCurl)

makeFlaredf <- function(i='01', type='b', 
                        fluxFileName = '01/sci_xrsf-l2-avg1m_g16_d20210101_v2-2-0.nc', 
                        flareFileName = '01/sci_xrsf-l2-flsum_g16_d20210101_v2-2-0.nc') {
  # Download 1-min flux info:
  fluxUrl0 = 'https://data.ngdc.noaa.gov/platforms/solar-space-observing-satellites/goes/goes16/l2/data/xrsf-l2-avg1m_science/2021/'
  fluxUrl = paste0(fluxUrl0,fluxFileName)
  if(url.exists(fluxUrl)) {
    fluxDestFile <- paste0("C:/Users/bklingen/Downloads/Flares/flux", i, ".nc")
    download.file(fluxUrl, fluxDestFile, method = 'auto', mode='wb')
    Flux0 = nc_open(fluxDestFile)
    dfFlux = tibble(
      secs_since_2000 = ncvar_get(Flux0, varid="time"),
      Date = as.POSIXct(secs_since_2000, origin = "2000-01-01 12:00:00"),
      Flux = ncvar_get(Flux0, varid=paste0("xrs",type,"_flux")),
    )
    nc_close(Flux0)
    #file.remove(fluxDestFile)
  } else {
    dfFlux = tibble(
      as.Date(character())
    )
  }
    
  # Download flare occurrence info:
  flareUrl0 = 'https://data.ngdc.noaa.gov/platforms/solar-space-observing-satellites/goes/goes16/l2/data/xrsf-l2-flsum_science/2021/'
  flareUrl = paste0(flareUrl0,flareFileName)
  if(url.exists(flareUrl)) {
    flareDestFile <- paste0("C:/Users/bklingen/Downloads/Flares/flare", i , ".nc")
    download.file(flareUrl, flareDestFile, method = 'auto', mode='wb')
    Flare0 = nc_open(flareDestFile)
    dfFlares = tibble(
      secs_since_2000 = ncvar_get(Flare0, varid="time"),
      Status = ncvar_get(Flare0, varid="status"),
    )
    nc_close(Flare0)
  } else {
    dfFlares = tibble(
      secs_since_2000 = integer(),
      Status = character(),
    )
  }
  ## joint dfFlux and dfFlare data:
  df = left_join( x = dfFlux,
                  y = dfFlares,
                  by = 'secs_since_2000')

  return(df %>% mutate(Flare = if_else(Status == 'EVENT_PEAK', 1, 0, missing = 0)))
}
```

Cycle over all 12 month in 2021:
```{r}
Flares2021 = tibble(
      secs_since_2000 = integer(),
      Date = as.Date(character()),
      Flux = numeric(),
      Status = character(),
      Flare = integer()
)

monthDays = c(31, 28, 31, 30, 31, 30, 31, 31, 30, 31, 30, 31)
monthNum = c('01', '02', '03', '04', '05', '06', '07', '08', '09', '10', '11', '12') 


for (k in 1:12) {
  days <- formatC(seq(1, monthDays[k]), width = 2, format = "d", flag = "0")

  for(i in days) {
    fluxFileName =  paste0(monthNum[k], '/sci_xrsf-l2-avg1m_g16_d2021', monthNum[k], i , '_v2-2-0.nc')
    flareFileName = paste0(monthNum[k], '/sci_xrsf-l2-flsum_g16_d2021', monthNum[k], i , '_v2-2-0.nc')
    df0 = makeFlaredf(i, type='b', fluxFileName=fluxFileName, flareFileName=flareFileName)
    Flares2021 = bind_rows(Flares2021,df0)
  }
}

save(Flares2021, file="C:/Users/bklingen/Downloads/Flares/Flares2021.r")

```

Now, create a matrix with 61 columns, where column nr. 31 is the time point at which a flare occurred, columns 1 - 30 are the values of the previous 30 fluxes (so, going back 30 minutes), and columns 32 - 61 are the values of the 30 fluxes after the declared flare:
```{r}
## Load in Flares2021 dataset
load(file="C:/Users/bklingen/Downloads/Flares/Flares2021.r")
flare_indices = which(Flares2021$Flare == 1)
numFlares = length(flare_indices)
FlareProfiles = matrix(NA, nrow=numFlares, ncol=63)
j = 1
for (i in flare_indices){
  if(i > 30 && i <= nrow(Flares2021) - 30) {
    FlareProfiles[j,] = c( ## standardize fluxes, equals 0 for flux at flar eevent
      (Flares2021$Flux[i] - Flares2021$Flux[(i-30) : (i + 30)]) / Flares2021$Flux[i],
      1,
      Flares2021$secs_since_2000[i]
    )
    j = j + 1
  }
}

## Controls: pick random index:
NotFlareProfiles = matrix(NA, nrow=50*numFlares, ncol=63)
nonflare_indices = setdiff(c(1:nrow(Flares2021)), flare_indices)
j = 1
for (i in sample(nonflare_indices, 50*numFlares)){
  if(i > 30 && i <= nrow(Flares2021) - 30) {
    NotFlareProfiles[j,] = c(
      (Flares2021$Flux[i] - Flares2021$Flux[(i-30) : (i + 30)]) / Flares2021$Flux[i],
      0,
      Flares2021$secs_since_2000[i]
    )
    j = j + 1
  }
}

library(tidyr)
library(dplyr)
library(readr)
Profiles = as_tibble(rbind(FlareProfiles, NotFlareProfiles))
dim(Profiles)
colnames(Profiles) = c(paste0('x', 1:61), 'y', 'SecsSince2000')

Profiles = Profiles %>% 
  drop_na() %>%  
  mutate(Date = as.POSIXct(SecsSince2000, origin = "2000-01-01 12:00:00")) %>% 
  select(-SecsSince2000)

ProfilesShuffeled = Profiles[sample(1:nrow(Profiles)),]
save(ProfilesShuffeled, file="C:/Users/bklingen/Downloads/Flares/ProfilesShuffeled.r")

======#write_csv(as.data.frame(ProfilesShuffeled) , file="C:/Users/bklingen/Downloads/Flares/ProfilesShuffeled.csv")
```
